Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-11 paper/paper.tex
Remove iterations  column from PLDI results table Made results tables \tiny
Added experimental discussion & stats breakdown

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}

@@

To illustrate some of these points, we were able to implement a first
version of {\sc kalashnikov} in around 150 lines of code, which took around
two hours to write.  The resulting code was easy to modify, optimise and
debug.  The current version, which includes support for floating-point
arithmetic, several optimisations and two model checking algorithms, is
still less than 1200 lines of code (600 lines of \newC and 600 lines of
Python).

Another benefit of using \newC as our logic is that it would be easy to
automatically extract specifications from C programs, which would be
useful in the context of superoptimisation.  This is because \newC is
a relatively large fragment of C, so we anticipate that many interesting
sections of real code will already be~\newC.

@@

To give a reference point for the efficiency of our tool, we present the results given for {\sc brahma}
on the same benchmarks, as reported in~\cite{brahma}.  These experiments were performed an 8-core,
1.86\,GHz Xeon with 4\,GB of RAM, so the timings should be comparable to within a factor of 2 or so.

The results of our experiments are shown in Fig.~\ref{fig:results-table}.  On the problems for which
we are able to synthesise a program, we take a similar time to {\sc brahma}.  We are unable to synthesise
programs for 6 of the 25 problems that {\sc brahma} is able to solve.  We believe this to be a result of
the slight difference in the problems solved by {\sc kalashnikov} and {\sc brahma}.  We do not ask the
user for any information beyond the program specification, whereas {\sc brahma} asks for an extended
component library when synthesis with the standard library fails.  This extended library contains a
reasonable amount of information about the structure of the final program, which we believe makes
problems P19 to P25 tractable for {\sc brahma} but intractable for us.

@@

\begin{figure}
\begin{center}
\include{figures/results_table}
\end{center}

\caption{Times for synthesis of machine integer benchmarks from {\sc brahma} PLDI paper}
\label{fig:results-table}
\end{figure}

\begin{figure}
\begin{center}
\include{figures/icse_results_table}
\end{center}

\caption{Times for synthesis of machine integer benchmarks from {\sc brahma} ICSE paper}
\label{fig:icse-results-table}
\end{figure}

@@

@@

>>>>

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsfonts}

@@

@@

To give a reference point for the efficiency of our tool, we present the results given for {\sc brahma}
on the same benchmarks, as reported in~\cite{brahma} and~\cite{brahma-icse}.  These experiments were performed an 8-core,
1.86\,GHz Xeon with 4\,GB of RAM.  Although there are 49 results shown across the two tables, it should
be noted that there is a lot of overlap between the two sets of benchmarks and that there are only
29 unique benchmarks altogether.

We also ran {\sc AHA} on our benchmarks and found that on the small problems (those with 2-instruction solutions)
{\sc AHA} found a single correct solution in negligible time.  On the larger benchmarks, candidate solutions
were still found rapidly (around 10\,s for the 3-instruction programs) but the number of candidates was
too large to practically check them all.  For example on p12, {\sc AHA} found 2284262 candidate
programs in 12.13\,s.

The results of our experiments are shown in Fig.~\ref{fig:results-table} and Fig.~\ref{fig:icse-results-table}.
The results include the runtime for each tool, the number of lines in the synthesised program and whether
{\sc Brahma} was able to complete the synthesis completely automatically (the Aut. column).
On both sets of benchmarks, the results can be broadly divided into three categories:

\paragraph{Problems where both {\sc Brahma} and {\sc Kalashnikov} synthesise minimal programs}
This represent the majority of cases.  A statistical breakdown of this case is shown in Fig~\ref{fig:stats-table}.
{\sc Kalashnikov} performs substantially better than {\sc Brahma} on these tests, and very similarly to
semibiased {\sc Brahma} (an optimisation introduced in~\cite{brahma-icse}).

\paragraph{Problems where {\sc Kalashnikov} is unable to synthesise a program}
This happens in 8 of the 29 cases.  In each case, {\sc Brahma} needs user guidance to synthesise the program.

\paragraph{Problems where {\sc Kalashnikov} synthesises a shorter program than {\sc Brahma}}
This happens in 4 of the 29 cases.  For example p24 in Fig.~\ref{fig:icse-results-table} uses
an obfuscated procedure

@@

\begin{figure}
\begin{center}
{\tiny
\include{figures/results_table}
}
\end{center}
\caption{Time for synthesis of machine integer benchmarks from~\cite{brahma}
 \label{fig:results-table}}
\end{figure}

\begin{figure}
\begin{center}
{\tiny
\include{figures/icse_results_table}
}
\end{center}
\caption{Times for synthesis of machine integer benchmarks from~\cite{brahma-icse}}
\label{fig:icse-results-table}
\end{figure}

@@

{\tiny
\begin{center}
 \begin{tabular}{r||r|r|r|r|r|r|r|r|r}
  & \multicolumn{3}{c}{Count} & \multicolumn{2}{|c}{Total time} & \multicolumn{2}{|c}{Absolute speedup} & \multicolumn{2}{|c}{Relative speedup} \\
  {\sc Brahma} version & Total & {\sc Kalashnikov} & {\sc Brahma} & {\sc Kalashnikov} & {\sc Brahma} & Mean & Median & Mean & Median \\
  &       & faster            & faster       &                   &              &      &        &      & \\
  \hline
  \hline
  PLDI & 16 & 13 & 3 & 184.12\,s & 544.20\,s & 22.50\,s & 2.11\,s & 2.60 & 1.77 \\
  ICSE Semibiased & 17 & 7 & 10 & 225.49\,s & 262.24\,s & 2.16\,s & -1.00\,s & 1.26 & 0.87
 \end{tabular}
\end{center}
}

 \caption{Performance statistics when {\sc Brahma} and {\sc Kalashnikov} were both successful
 \label{fig:stats-table}}
\end{figure}

@@

\paragraph{Future Work}
Another benefit of using \newC as our logic is that it would be easy to
automatically extract specifications from C programs, which would be
useful in the context of superoptimisation.  This is because \newC is
a relatively large fragment of C, so we anticipate that many interesting
sections of real code will already be~\newC.
